---
title: "Rasch in R and Quarto, part 1 (updated)"
subtitle: "Intro and WAAQ-7"
title-block-banner: "#009ca6"
author: 
  name: Magnus Johansson
  affiliation: RISE Research Institutes of Sweden
  affiliation-url: https://www.ri.se/sv/vad-vi-gor/expertiser/kategoriskt-baserade-matningar
  orcid: 0000-0003-1669-592X
date: '2022-08-16'
always_allow_html: true
format: 
  html:
    toc: true
    self-contained: true
    logo: rise_logo_quarto.png
    mainfont: 'Lato'
    monofont: 'Roboto Mono'
    code-overflow: wrap
    code-tools: true
    code-fold: true
    number-sections: true
    fig-dpi: 300
    layout-align: left
  pdf:
    papersize: a4
    documentclass: article #article, report or book
    classoption: [twocolumn, portrait]
  revealjs:
    theme: default
    logo: rise_logo_quarto.png
    chalkboard: false
    self-contained: true
    footer: 'Material skapat av magnus.p.johansson@ri.se'
    mainfont: 'Lato'
    slide-level: 4
    scrollable: true
    smaller: false
  docx:
    toc: true
    number-sections: true
    title: "WAAQ-7"
    subtitle: "Rasch analysis"
    author: "Magnus Johansson"
#    reference-doc: RISEmallMJv6.dotx
execute:
  echo: false
  warning: false
  message: false
  cache: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: inline
bibliography: 
- references.bib
- grateful-refs.bib
---

```{r}
#| label: blurb
#| include: false

### Please check for the latest version of this file at:
### https://github.com/pgmj/pgmj/blob/main/RaschR1v2.qmd 
### and the companion file with more code chunks:
### https://github.com/pgmj/pgmj/blob/main/RaschR2v2.qmd

### The R package RISEkbmRasch needs to be installed using
# install.packages('devtools')
# devtools::install_github("pgmj/RISEkbmRasch")
# package grateful is installed the same way:
# devtools::install_github("Pakillo/grateful")

```

## Why Rasch analysis in R?

::: incremental
-   Traceability
    -   Transparency in decisionmaking
-   Reproducibility
-   Freely available
-   Documentation in the same file as analysis
    -   (with Quarto/Rmarkdown)
:::

### Agenda overview

::: incremental
-   Sample analysis with provided data and code
    -   WAAQ-7
    -   PSS-10
-   Suggested structure for analysis
    -   Confirmatory and/or exploratory
:::

### Quarto formatting basics

<https://quarto.org/docs/authoring/markdown-basics.html>

-   Keep the page above open in a web browser
-   We will use library(kableExtra) for tables
    -   since it works with HTML/revealjs and PDF output
    -   even though library(formattable) looks nicer...

### Exercise 1

::: incremental
-   I'll walk you through the Rstudio interface
-   You'll run the setup code chunk, and practice navigation a bit
    -   Enable/disable *Outline*
    -   Navigate with Outline
    -   Navigate with *chunks*
    -   Switch between *source* and *visual* editor
    -   Changing size of the quadrant windows
    -   Look at the objects *df* and *itemlabels*
:::

### Exercise 2

-   We will walk through the code chunks together
-   Learn what they do, but not how (in detail)
-   I'll explain some basic R concepts as we move ahead
-   Focus on how you can structure the analysis
-   Discuss the output briefly
    -   feedback on how to improve is welcome :)
-   Small exercises are described at the end of the file

### Exercise 3 (later on)

-   In the folder "Exercise2" you will find a RaschR2 qmd file
-   It contains a setup block that will import a dataset (PSS-10)
-   You can practice using this file (RaschR1) as a template
    -   copy code chunks for your analysis from this file to the other
        file
    -   copy code, run the chunks, check the output, write commentary.
        **Repeat**
    -   *there is a folder called "Exercise2solution" that shows how I did
        the analysis*

```{r}
#| label: setup
#| code-fold: false
#| include: false

# two packages below require that you use devtools to install them manually:
# first install devtools by
# install.packages('devtools')

library(car)
library(grateful) # devtools::install_github("Pakillo/grateful")
library(kableExtra)
library(readxl)
library(tidyverse)
library(eRm)
library(ggrepel)
library(mirt)
library(psych)
library(ggplot2)
library(psychotree)
library(matrixStats)
library(reshape)
library(knitr)
library(cowplot)
library(formattable) 
library(RISEkbmRasch) # devtools::install_github("pgmj/RISEkbmRasch")
library(HH)

### some commands exist in multiple packages, here we define preferred ones that are frequently used
select <- dplyr::select
count <- dplyr::count
recode <- car::recode
rename <- dplyr::rename

### set up color palette based on RISE guidelines
RISEprimGreen <- "#009ca6"
RISEprimRed <- "#e83c63"
RISEprimYellow <- "#ffe500"
RISEprimGreenMid <- "#8dc8c7"
RISEprimRedMid <- "#f5a9ab"
RISEprimYellowMid <- "#ffee8d"
RISEprimGreenLight <- "#ebf5f0"
RISEprimRedLight <- "#fde8df"
RISEprimYellowLight <- "#fff7dd"
RISEcompPurple <- "#482d55"
RISEcompGreenDark <- "#0e4e65"
RISEgrey1 <- "#f0f0f0"
RISEgrey2 <- "#c8c8c8"
RISEgrey3 <- "#828282"
RISEgrey4 <- "#555555"

# set some colors used later
cutoff_line <- RISEprimRed
dot_color <- "black"
backg_color <- RISEprimGreenLight

# set fontsize for all tables
r.fontsize <- 15

### first we pre-set our chosen cut-off values for some commonly used indices:
msq_min <- 0.7
msq_max <- 1.3
zstd_min <- -2
zstd_max <- 2
loc_dep <- 0.2 # above average residual correlation
dif_dif <- 0.5 # logits difference between groups in average item location (DIF)

### zstd is inflated with large samples (N > 500). Reduce sample size to jz and 
### run analysis yz random samples to get average ZSTD
jz = 300 # number to include in dataset
yz = 10 # number of random samples

# Load data
df<-as.data.frame(read_excel("C:/Users/magnuspjo/RISE/KP Centrum för Kategoriskt Baserade Mätningar - Dokument/Kunskapsutveckling/RaschR/data/GNI23data v1.1.xls"))

### if you need to download the datafile:
# library(readODS)
# url <- "https://figshare.com/ndownloader/files/27953790"
# destfile <- "gnidata.ods"
# download.file(url, destfile, mode="wb")
# df <- read_ods(file = "gnidata.ods")

# Load item information
itemlabels<-read_excel("C:/Users/magnuspjo/RISE/KP Centrum för Kategoriskt Baserade Mätningar - Dokument/Kunskapsutveckling/RaschR/data/WAAQitemlabels.xls")
#responseOptions<-read_excel("C:/Users/magnuspjo/RISE/KP Centrum för Kategoriskt Baserade Mätningar - Dokument/Kunskapsutveckling/RaschR/data/PSS10itemlabels.xls", sheet = 2)
#introText<-read_excel("C:/Users/magnuspjo/RISE/KP Centrum för Kategoriskt Baserade Mätningar - Dokument/Kunskapsutveckling/RaschR/data/PSS10itemlabels.xls", sheet = 3)

# our datafile contains miscoded variable, the code below makes all variables numeric
# this is not needed/desirable for all datasets, so use with care
df <- df %>% 
  mutate_if(is.character, as.numeric)

# Make a backup of the dataframe
df.all<-df

# Create a vector with variables names
scale.items <- itemlabels$itemnr

# recode WAAQ items to 0-6 (instead of 1-7)
for (i in scale.items) {
  df[[i]]<-recode(df[[i]],"1=0;2=1;3=2;4=3;5=4;6=5;7=6",as.factor=FALSE)
}

##### Code for dealing with missing data

# Create a figure showing % of missing data based on the complete dataset
missingdata <- df %>% 
  select(starts_with("WAAQ")) %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(Missing = rowSums(is.na(.))) %>% 
  select(Missing) %>% 
  arrange(desc(Missing)) %>% 
  rownames_to_column(var = "Item") %>% 
  mutate(Percentage = Missing/nrow(df)*100) %>% 
  mutate(Item = factor(Item, levels = rev(Item))) %>%
  ggplot(aes(x = Item, y = Percentage)) +
  geom_col() +
  coord_flip() +
  ggtitle("Missing data per item") +
  xlab("Items") +
  ylab("Percentage of responses missing")

# Minimum items responses that a participant should have to be included in the analysis?
min.responses <- 3

# Select the variables we will work with, and filter out respondents with a lot of missing data
df.omit.na <- df %>% 
  select(starts_with("WAAQ"),Sex,Age) %>%  # variables that start with "WAAQ"
  filter(length(scale.items)-rowSums(is.na(.[scale.items])) >= min.responses) # include only respondents with data for at least 3 items

###
  
# DIF variables into vectors
dif.gender <- df.omit.na$Sex
dif.age <- df.omit.na$Age
# then remove them from dataframe, since we need a dataframe with only item data for the Rasch analyses
df.omit.na$Sex <- NULL
df.omit.na$Age <- NULL

# set item names to neutral qX (and avoid issues with the character "_" in pdf output)
#names(df.omit.na) <- paste0("q", c(1:ncol(df.omit.na)))

```

## Background information {.smaller}

The Work-related Acceptance and Action Questionnaire [@bond2013] is a
measure of psychological flexibility contextualized to a workplace
setting.

The WAAQ uses 7 questions, all using the same response scale with 7
steps (ranging from *"Stämmer aldrig"* to *"Stämmer alltid"*, with
labels on all categories).

```{r}
#| label: showitems
#| tbl-cap: "Items in the Swedish WAAQ"
#| tbl-cap-location: top

RIlistitems(df.omit.na)
  
```

### Participants {.smaller}

We have `r nrow(df.omit.na)` participants and their gender distribution
is shown in the table below. All participants are Swedish adults and the
dataset was collected to validate a measure of social interactions at
workplaces [@johansson2021; @johansson2021a].

```{r}
#| label: genderdistr
#| tbl-cap: "Gender distribution in sample"
#| tbl-cap-location: top

# Make table object to show gender counts and percentages
dif.gender %>%
  recode("1='Male';2='Female'") %>% # change these values depending on demographic variable
  table() %>% 
  as_tibble() %>% 
  mutate('Percent' = (round((100 * n / sum(n)),1))) %>% 
  dplyr::rename('Gender' = '.') %>% 
  kbl(booktabs = T, escape = F, table.attr = "style='width:20%;'") %>%
  # options for HTML output
  kable_styling(bootstrap_options = c("striped", "hover"), 
                position = "center",
                full_width = T,
                font_size = r.fontsize,
                fixed_thead = T) %>% 
  column_spec(1, bold = T) %>% 
  kable_classic(html_font = "Lato") %>% 
  # latex_options are for PDF output
  kable_styling(latex_options = c("striped","scale_down"))

## Alternative option:
#RIdemographics(dif.gender, "Gender")
# this may need explanatory text, as the function does not include any recoding of numerical demographic categories
                
```

### Descriptives of raw data

Responses to all items are summarized below.

```{r}
#| label: desc1
#| tbl-cap: "Total number of responses for all items"
RIallresp(df.omit.na)
```

### Descriptives - item level

::: panel-tabset
### Tile plot
```{r}
#| label: desc1.2
RItileplot(df.omit.na)
```
### Stacked bars
```{r}
#| label: desc1.3
RIbarstack(df.omit.na)
```
### Barplots
```{r}
#| label: desc1.5
#| layout-ncol: 2
RIbarplot(df.omit.na)
```
:::

#### Recoding responses {.smaller}

There are fewer than 10 responses in category 0 for all items, which is
problematic for the analysis. We will merge the two lowest categories.
For item 2 the four lowest categories will be merged.

```{r}
#| label: recode

# recode all variables in dataframe (from column 1 to the maximum number of columns)
for (i in 1:ncol(df.omit.na)) {
  df.omit.na[,i]<-recode(df.omit.na[,i],"1=0;2=1;3=2;4=3;5=4;6=5",as.factor=FALSE)
}
# then recode a single variable, q2
df.omit.na$WAAQ_2<-recode(df.omit.na$WAAQ_2,"1=0;2=0;3=1;4=2;5=3",as.factor=FALSE)

# make a new tile plot to review results
RItileplot(df.omit.na)

```

Some cells still have fewer than 10 responses. Let's look at the response categories using Rasch modeling to produce ICC plots.

## Analysis of response categories {.smaller}

```{r}
#| label: respcat1
#| include: false

# package mirt is used to create a compact figure with all item ICCs

mirt.rasch <- mirt(df.omit.na, model=1, itemtype='Rasch') # unidimensional Rasch model
```

```{r}
#| label: respcatfig1
plot(mirt.rasch, type="trace")
# for bigger, better looking, individual figures:
#df.erm<-PCM(df.omit.na) # run PCM model, replace with RSM (rating scale) or RM (dichotomous) for other models
#plotICC(df.erm, xlim = c(-6, 8), legpos = FALSE, ylab = "Sannolikhet", xlab = "Övergripande förmåga")

```

Items 1 and 5 have issues with disordered response categories, which
need to be addressed. For both items the two lowest categories will be
merged.

### ICC plots for recoded items

```{r}
#| label: recode2
#| layout-ncol: 2

df.omit.na$WAAQ_1<-recode(df.omit.na$WAAQ_1,"1=0;2=1;3=2;4=3;5=4",as.factor=FALSE)
df.omit.na$WAAQ_5<-recode(df.omit.na$WAAQ_5,"1=0;2=1;3=2;4=3;5=4",as.factor=FALSE)

# individual plots for those two items:
RIitemcats(df.omit.na, items = c("WAAQ_1","WAAQ_5"))

```

Response categories are working as expected, after the adjustments made.

### Floor/ceiling effects - raw data {.smaller}

```{r}
#| label: rawdist2

RIrawdist(df.omit.na)

```

## Rasch analysis 1 {.smaller}

The eRm package, which uses Conditional Maximum Likelihood (CML)
estimation, will be used primarily. For this analysis, the Partial
Credit Model will be used.

::: panel-tabset
### Item fit {.smaller}
```{r}
#| label: rasch1.fit
RIitemfitPCM(df.omit.na)
```
### PCA
```{r}
#| label: rasch1.pca
#| tbl-cap: "PCA of Rasch model residuals"
RIpcmPCA(df.omit.na)
```
## 1st contrast loadings
```{r}
#| label: rasch1.load
RIloadLoc(df.omit.na)
```
### Residual correlations {.smaller}
```{r}
#| label: rasch1.rcorr
RIresidcorr(df.omit.na, cutoff = 0.2)
```
### Targeting
```{r}
#| label: rasch1.targ
RItargeting(df.omit.na)
```
### Targeting sorted
```{r}
#| label: rasch1.targs
df.erm<-PCM(df.omit.na)
plotPImap(df.erm, sorted = T)
```
### Reliability
```{r}
#| label: rasch1.rel
RItif(df.omit.na)
```
### Items
```{r}
#| label: rasch1.items
RIlistitems(df.omit.na)
```
:::

Values in red are below/above the pre-set cutoff values.

A correlation matrix is created based on the Rasch model residuals for
each item. The average correlation is calculated, and item pairs that
correlate more than the pre-set cutoff value (see footnote) above the
average correlation are indicated in red in the table below.

Items 1 and 3 are correlated above the relative cutoff value. 
Looking at the item descriptions, both are related to
working efficiently when being worried in general (item 3) or being
worried about personal problems. As we saw in the section on item fit,
item 3 had quite low item fit values. This is a common issue with broad
questions.

It would be reasonable to eliminate either question 1 or 3, since they
clearly overlap in content. We will leave this for an exercise.

## DIF analysis {.smaller}

Our sample is too small to conduct a properly powered DIF-analysis of any of
the demographic variables available (gender/age). N \> 250 per subgroup
would be ideal, and N \> 150 minimal.

For the sake of practice, we will go through the steps in setting up and
conducting a DIF analysis anyway.

We will use the Rasch Tree tool [@strobl2013], which makes it possible
to evaluate interaction DIF, if sample size allows. For this analysis,
we will only investigate gender effects on DIF.

```{r}
#| label: dif1

RIdifTable(df.omit.na, dif.var = dif.gender)
# if no DIF is found, only a text saying so will be in the output
# if DIF is found, a figure and a table is created
# RIdifFigure() can then be used, if desired
# same functions ending with RM are available for dichotomous data

```

No DIF was found in our small sample.

## Targeting & item/person locations

```{r}
#| label: wrightm1
#| fig-width: 8
#| fig-height: 8

RItargeting(df.omit.na) # use option dich = TRUE for dichotomous data

# another option is:
df.erm <- PCM(df.omit.na)
plotPImap(df.erm, sorted = TRUE) # sorts items based on location (black dots)

```

### Reliability {.smaller}

"Test information" shows the collective information based on the
combination of all items into a unidimensional scale.

```{r}
#| label: tif1

RItif(df.omit.na)

# if you want other limits for the x axis:
#RItif(df.omit.na, lo = -8, hi = 8)

```

The horizontal lines at TIF 3.33 and 5.0 indicate the conventional 
cutoff values for acceptable reliability of 0.7 and 0.8, respectively.

The table footnote includes information on floor/ceiling effects.

#### Item information {.smaller}

"Item information" shows individual curves indicating the amount of
information contributed by each item.

```{r}
#| label: iteminf
df.erm <- PCM(df.omit.na)
plotINFO(df.erm, type = "item", legpos = "topleft") # get item information curves
```

### Person location and infit ZSTD {.smaller}

```{r}
#| label: personfit1

RIpfit(df.omit.na)

```

## Item parameters

```{r}
#| label: itemparams

RIitemparams(df.omit.na)

```

## Ordinal to interval transformation table

```{r}
#| label: ordintTr
#| include: false
df.erm<-PCM(df.omit.na) # run PCM model, replace with RSM (rating scale) or RM (dichotomous) for other models
# get estimates, code borrowed from https://bookdown.org/chua/new_rasch_demo2/PC-model.html
item.estimates <- eRm::thresholds(df.erm)
item_difficulty <- item.estimates[["threshtable"]][["1"]]
item_difficulty<-as.data.frame(item_difficulty)
item.se <- item.estimates$se.thresh
person.locations.estimate <- person.parameter(df.erm) 
ple<-as.data.frame(print(person.locations.estimate))
rownames(ple)<-NULL
colnames(ple)<-c("Ordinal sum score","Logits","SE (Logits)")
ple[,2]<-round(ple[,2],2)
ple[,3]<-round(ple[,3],2)

ple$'Interval score (0-100)'<-round(scales::rescale(ple[,2], to = c(0,100)),1) # set 0,100 to desired range
ple$'SE (0-100)'<-round((ple$`SE (Logits)`/(0.83/8)),1) # approximated from looking at the table
ple$'MU (0-100)'<-(ple$'SE (0-100)')*2
```

```{r}
#| label: ordintTbl
#| tbl-cap: "Ordinal-interval transformation table"
#| layout-ncol: 2

ple %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  kbl(booktabs = T, escape = F) %>%
  # bootstrap options are for HTML output
  kable_styling(bootstrap_options = c("striped", "hover"), 
                position = "left",
                full_width = F,
                font_size = r.fontsize,
                fixed_thead = T) %>% # when there is a long list in the table
#  column_spec(c(2:3), color = "red") %>% 
#  row_spec(3:5, bold = T, color = "white", background = "lightblue") %>% 
  column_spec(1, bold = T) %>% 
  kable_classic(html_font = "Lato") %>% 
  # latex_options are for PDF output
  kable_styling(latex_options = c("striped","scale_down")) %>% 
  footnote(general = "SE = Standard Error. Multiply by 2 for MU = Measurement Uncertainty")
```

## Item location & infit ZSTD

```{r}
#| label: locinfz

RIinfitLoc(df.omit.na)

```

## Item location & outfit ZSTD

```{r}
#| label: locoutfz

RIoutfitLoc(df.omit.na)

```

## First contrast loadings & item locations

```{r}
#| label: locload

RIloadLoc(df.omit.na)

```

The figure above shows the item loadings on the first contrast factor of the PCA of Rasch model residuals, and the item locations.

## Software used

```{r}
#| label: packagesv
pkgs <- cite_packages(cite.tidyverse = TRUE, 
                      output = "table",
                      bib.file = "grateful-refs.bib",
                      include.RStudio = TRUE)
formattable(pkgs, 
            table.attr = 'class=\"table table-striped\" style="font-size: 15px; font-family: Lato; width: 80%"')
```

## Exercises

1.  Make a table of partipant age distribution
2.  Eliminate either item 1 or 3 and check the effects on reliability
    - hint: df.omit.na$item <- NULL
3.  Bonus if you are quick:
    -   Run a DIF analysis using age as DIF variable
    -   first add the dif.age variable in the setup chunk
    -   then duplicate the DIF chunk and modify code

## References
