---
title: "Rasch in R and Quarto, part 2"
subtitle: "PSS-10"
title-block-banner: "#009ca6"
author: 
  name: Magnus Johansson
  affiliation: RISE Research Institutes of Sweden
  affiliation-url: https://ri.se/shic
  orcid: 0000-0003-1669-592X
date: '2022-06-13'
format: 
  html:
    toc: true
    self-contained: true
    logo: rise_logo_quarto.png
    mainfont: 'Lato'
    monofont: 'Roboto Mono'
    code-overflow: wrap
    code-tools: true
    code-fold: true
    number-sections: true
    fig-dpi: 300
  pdf:
    papersize: a4
    documentclass: report #article, report or book
    classoption: [onecolumn, portrait]
  revealjs:
    theme: default
    logo: rise_logo_quarto.png
    chalkboard: true
    self-contained: false
#    footer: 'Material skapat av magnus.p.johansson@ri.se'
    mainfont: 'Lato'
    slide-level: 4
    scrollable: true
    smaller: true
execute:
  echo: false
  warning: false
  message: false
  cache: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: inline
bibliography: references.bib
---

```{r}
#| label: setup
#| code-fold: false
#| include: false

library(car)
library(grateful)
library(kableExtra)
library(readxl)
library(tidyverse)
library(eRm)
library(ggrepel)
library(mirt)
library(psych)
library(ggplot2)
library(psychotree)
library(matrixStats)
library(reshape)
library(knitr)
library(cowplot)
library(formattable)
library(RISEkbmRasch)
library(HH)

### some commands exist in multiple packages, here we define preferred ones that are frequently used
select <- dplyr::select
count <- dplyr::count
recode <- car::recode
rename <- dplyr::rename

### set up color palette based on RISE guidelines
RISEprimGreen <- "#009ca6"
RISEprimRed <- "#e83c63"
RISEprimYellow <- "#ffe500"
RISEprimGreenMid <- "#8dc8c7"
RISEprimRedMid <- "#f5a9ab"
RISEprimYellowMid <- "#ffee8d"
RISEprimGreenLight <- "#ebf5f0"
RISEprimRedLight <- "#fde8df"
RISEprimYellowLight <- "#fff7dd"
RISEcompPurple <- "#482d55"
RISEcompGreenDark <- "#0e4e65"
RISEgrey1 <- "#f0f0f0"
RISEgrey2 <- "#c8c8c8"
RISEgrey3 <- "#828282"
RISEgrey4 <- "#555555"

# set some colors used later
cutoff_line <- RISEprimRed
dot_color <- "black"
backg_color <- RISEprimGreenLight

### first we pre-set our chosen cut-off values for some commonly used indices:
msq_min <- 0.7
msq_max <- 1.3
zstd_min <- -2
zstd_max <- 2
loc_dep <- 0.2 # above average residual correlation
dif_dif <- 0.5 # logits difference between groups in average item location (DIF)

### zstd is inflated with large samples (N > 500). Reduce sample size to jz and 
### run analysis yz random samples to get average ZSTD
jz = 300 # number to include in dataset
yz = 10 # number of random samples

# Load data
df<-as.data.frame(read_excel("C:/Users/magnuspjo/RISE/KP Centrum för Kategoriskt Baserade Mätningar - Dokument/Kunskapsutveckling/RaschR/data/GNI23data v1.1.xls"))

### if you need to download the datafile:
# library(readODS)
# url <- "https://figshare.com/ndownloader/files/27953790"
# destfile <- "gnidata.ods"
# download.file(url, destfile, mode="wb")
# read_ods(file = "gnidata.ods")

df <- df %>% 
  mutate_if(is.character, as.numeric)

# Make a backup of the dataframe
df.all<-df

# Select the variables we will work with, which as this point can only be items
df.omit.na <- df %>% 
  select(starts_with("PSS"),Sex,Age) %>% # variables that start with "WAAQ"
  na.omit() # remove rows with 100% missing data

# DIF variables into vectors
dif.gender <- df.omit.na$Sex
dif.age <- df.omit.na$Age
# them remove them from dataframe
df.omit.na$Sex <- NULL
df.omit.na$Age <- NULL

# set item names to neutral qX (and avoid issues with _ in pdf output)
names(df.omit.na) <- paste0("q", c(1:ncol(df.omit.na)))

# recode items to 0-6 (instead of 1-7)
# for (i in 1:7) {
#   df.omit.na[,i]<-recode(df.omit.na[,i],"1=0;2=1;3=2;4=3;5=4;6=5;7=6",as.factor=FALSE)
# }

# Load item information
itemlabels<-read_excel("C:/Users/magnuspjo/RISE/KP Centrum för Kategoriskt Baserade Mätningar - Dokument/Kunskapsutveckling/RaschR/data/PSS10itemlabels.xls")
responseOptions<-read_excel("C:/Users/magnuspjo/RISE/KP Centrum för Kategoriskt Baserade Mätningar - Dokument/Kunskapsutveckling/RaschR/data/PSS10itemlabels.xls", sheet = 2)
introText<-read_excel("C:/Users/magnuspjo/RISE/KP Centrum för Kategoriskt Baserade Mätningar - Dokument/Kunskapsutveckling/RaschR/data/PSS10itemlabels.xls", sheet = 3)
```

## Background information

The Perceived Stress Scale [@lee2012; @nordin2013] is a measure of
subjective stress. It is available in three versions, with 14, 10 or 4
items. This dataset consists of the 10 item version, PSS-10.

All items have the same response options,
**`r responseOptions$ResponseText`**, which are recoded to 0-4 for this
analysis. The questions are introduced with the text
***`r introText$introtext`***.

The table below lists all items, with the four positively worded items
highlighted with green backgrounds.

```{r}
#| label: showitems
#| tbl-cap: "Items in the Swedish PSS-10"
#| tbl-cap-location: top

RIcolorlistitems(c(4,5,7,8), color = RISEprimGreenLight)

```

### Participants

We have `r nrow(df.omit.na)` participants and their gender distribution
is shown in the table below. All participants are Swedish adults and the
dataset was collected to validate a measure of social interactions at
workplaces [@johansson2021; @johansson2021a].

```{r}
#| label: genderdistr
#| tbl-cap: "Gender and age distribution in sample"
#| tbl-cap-location: top
#| layout-ncol: 2

# Make table object to show gender counts and percentages
dif.gender %>%
  recode("1='Male';2='Female'") %>% 
  table() %>% 
  as_tibble() %>% 
  mutate('Percent' = (round((100 * n / sum(n)),1))) %>% 
  dplyr::rename('Gender' = '.') %>% 
  kbl(booktabs = T, escape = F, table.attr = "style='width:20%;'") %>%
  # options for HTML output
  kable_styling(bootstrap_options = c("striped", "hover"), 
                position = "center",
                full_width = T,
                font_size = 12,
                fixed_thead = T) %>% 
  column_spec(1, bold = T) %>% 
  kable_classic(html_font = "Lato") %>% 
  # latex_options are for PDF output
  kable_styling(latex_options = c("striped"))

dif.age %>%
  recode("2='18-29';3='30-39';4='40-49';5='50-59';6='60+'") %>% 
  table() %>% 
  as_tibble() %>% 
  mutate('Procent' = (round((100 * n / sum(n)),1))) %>% 
  dplyr::rename('Age group' = '.') %>% 
  kbl(caption = "Age distribution", booktabs = T, escape = F, table.attr = "style='width:35%;'") %>%
  # options for HTML output
  kable_styling(bootstrap_options = c("striped", "hover"), 
                position = "center",
                full_width = T,
                font_size = 12,
                fixed_thead = T) %>% 
  column_spec(1, bold = T) %>% 
  kable_classic(html_font = "Lato") %>% 
  # latex_options are for PDF output
  kable_styling(latex_options = c("striped"))
                
```

Demographic subgroups are too small for properly powered DIF-analyses,
which should have at least 150 in each subgroup, and ideally 250 or
more.

## Descriptives of raw data

Responses to all items are summarized below.

```{r}
#| label: descriptives1
#| tbl-cap: "Total number of responses for all items"

RIallresp(df.omit.na)

```

## Descriptives - item level

::: panel-tabset
### Tile plot

```{r}
#| label: descriptives2

RItileplot(df.omit.na)

```

### Stacked bars

```{r}
#| label: stack1

RIbarstack(df.omit.na)
```

### Diverging bars

```{r}
#| label: diverg1

RIbardiv(df.omit.na)
```

### Barplots {.scrollable}

```{r}
#| label: alt-descriptives
#| layout-ncol: 2

RIbarplot(df.omit.na)

```
:::

#### Recoding responses

Since there are fewer than 10 responses in category 4 for most items we
will merge category 4 with category 3 before proceeding with the
analysis.

```{r}
#| label: recode

for (i in 1:ncol(df.omit.na)) {
  df.omit.na[,i]<-recode(df.omit.na[,i],"4=3",as.factor=FALSE)
}

RItileplot(df.omit.na)
RIbardiv(df.omit.na)

```

## Analysis of response categories

```{r}
#| label: respcat1
#| include: false

mirt.rasch <- mirt(df.omit.na, model=1, itemtype='Rasch') # unidimensional Rasch model
```

```{r}
#| label: respcatfig1
plot(mirt.rasch, type="trace")

```

### ICC plots for recoded items

```{r}
#| label: recode2
#| code-fold: true
#| layout-ncol: 2

df.omit.na$q4<-recode(df.omit.na$q4,"2=1;3=2",as.factor=FALSE)

# individual plots for those two items:
RIitemcats(df.omit.na, items = "q4")

```

Response categories are working as expected, after the adjustments made.

## Dimensionality

The eRm package, which uses Conditional Maximum Likelihood (CML)
estimation, will be used primarily. For this analysis, the Partial
Credit Model will be used.

First, we look at the eigenvalues from PCA analysis of Rasch residuals.
These should be below 2.0 to support unidimensionality.

```{r}
#| label: dim1
#| tbl-cap: "PCA of Rasch model residuals"

RIpcmPCA(df.omit.na)
```

The first eigenvalue exceeds 2.0, which indicates multidimensionality.

```{r}
#| label: modelfit1

### zstd is inflated with large samples (N > 500). Reduce sample size to jz and 
### run analysis yz random samples to get average ZSTD
RIitemfitPCM(df.omit.na, 300, 10)

```

Values in red are beyond the pre-set cutoff values. We'll leave them for
now.

### Residual correlations

A correlation matrix is created based on the Rasch model residuals for
each item.

The average correlation is calculated, and item pairs that correlate
more than the pre-set cutoff value of `r loc_dep` above the average
correlation are indicated in red in the table below.

```{r}
#| label: locdeps1

RIresidcorr(df.omit.na, cutoff = 0.2)
```

We have multiple item pairs with too large correlations:

-   q4 with 5 and 7
-   q7 with 8
-   q1 with 9

This means that all four positively worded items are problematic. They
will be removed, and a new analysis of item fit and residual
correlations conducted.

### Item fit 2, without positive items

```{r}
#| label: dim2
#| tbl-cap: "PCA of Rasch model residuals"

# create vector with eliminated items
removed_items<-c("q4","q5","q7","q8")

# select all items except those removed
df.omit.na <- df.omit.na %>% 
  select(!all_of(removed_items))

RIpcmPCA(df.omit.na)
``` 

Eigenvalues look good, supporting unidimensionality.

```{r}
#| label: modelfit2

RIitemfitPCM(df.omit.na)

RIlistitems(df.omit.na)

```

Item fit is acceptable for all items.

### Residual correlations 2

```{r}
#| label: locdeps2

RIresidcorr(df.omit.na, 0.2)

```

Item 1 and 9 are still correlated above the cutoff. Both items contain wording about being
upset/angry about things that are unexpected or outside of one's
control.

Both items have similar item fit. Let's look at the targeting properties
of these items to see how they compare.

### Targeting

```{r}
#| label: wrightm1

RItargeting(df.omit.na)
df.erm <- PCM(df.omit.na)
plotPImap(df.erm, sorted = TRUE)
```

Items 1 and 9 are virtually identical in threshold locations, let's see
how the DIF analysis works out before removing either of them.

## DIF analysis of gender

```{r}
#| label: dif1

RIdifTable(df.omit.na, dif.gender)

```

No significant gender DIF is indicated.

### DIF analysis of age

We will look at the age variable, for practice.

```{r}
#| label: dif2

RIdifTable(df.omit.na, dif.age)

```

Finally we get to investigate DIF size! One interesting property of the
Rasch Tree DIF method is that when a DIF variable has more than 2
levels, it will determine where the largest differences are, and build a
sequential tree if multiple significant differences are found.

In our sample, there was DIF of age found between those with category 2
or lower (age 18-29), and those with category 3 and above (age 30+).

The DIF table highlights values beyond 0.5 logits maximum difference between groups. If more than two groups are included in the DIF analysis, the lowest and highest value is automatically identified.

#### DIF size figure

```{r}
#| label: DIFsize

RIdifFigure(df.omit.na, dif.age)

```

Significant DIF size was found for items 3 and 9. This could be a good reason
to remove item 9 instead of item 1 (residual correlation problems
earlier). Our sample is too small to draw conclusions about DIF, but since
one of them needs to be removed, item 9 will be removed before moving on
to the Targeting section.

### DIF interaction (age + gender)

There is another unique function in the Rasch Tree DIF, which is the
interaction of multiple DIF variables. We can input age and gender at
the same time! Our groups will be way too small, so this is really only
to illustrate how to do it.

```{r}
#| label: dif3
#| echo: true

# set up the DIF analysis
df.tree <- data.frame(matrix(ncol = 0, nrow = nrow(df.omit.na))) # we need to make a new dataframe
df.tree$difdata <- as.matrix(df.omit.na) # containing item data in a nested dataframe
# and DIF variables:
df.tree$age<-dif.age
df.tree$gender<-dif.gender
pctree.out<-pctree(difdata ~ age + gender, data = df.tree)
plot(pctree.out)

```

No interaction DIF found (only age shows problems), which unfortunately means we won't get
to make a neat table showing multiple groups, item locations, means and
standard deviations.

## Targeting

```{r}
#| label: itemremoval
#| include: false

# delete item
df.omit.na$q9 <- NULL
# add to list of removed items
removed_items <- c(removed_items,"q9")

RItargeting(df.omit.na)
```

### Reliability

"Test/scale information" shows the collective information based on the
combination of all items into a unidimensional scale.

```{r}
#| label: tif1

RItif(df.omit.na)
```
Reliability is insufficient for the 5 item PSS scale.

#### Item information

"Item information" shows individual curves indicating the amount of
information contributed by each item.

```{r}
#| label: iteminf
df.erm <- PCM(df.omit.na)
plotINFO(df.erm, type = "item", legpos = "topleft") # get item information curves
```

### Person fit

```{r}
#| label: personfit1

RIpfit(df.omit.na)

```


## Item parameters

```{r}
#| label: itemparams
#| tbl-cap: "Item thresholds and locations"

RIitemparams(df.omit.na)
```

## Ordinal to interval transformation table

Since the scale is not reliable enough, this will not be provided.

## Software used

```{r}
#| label: packagesv


pkgs <- cite_packages(cite.tidyverse = TRUE, 
                      output = "table",
                      bib.file = "grateful-refs.bib",
                      include.RStudio = TRUE)

formattable(pkgs, 
            table.attr = 'class=\"table table-striped\" style="font-size: 15px; font-family: Lato; width: 80%"')
```

## References
