---
title: "Rasch in R and Quarto"
subtitle: "Template file"
title-block-banner: "#009ca6"
author: 
  name: Magnus Johansson
  affiliation: RISE Research Institutes of Sweden
  affiliation-url: https://www.ri.se/en/what-we-do/expertises/categorically-based-measurements
  orcid: 0000-0003-1669-592X
date: '2022-10-24'
always_allow_html: true
format: 
  html:
    toc: true
    self-contained: true
    logo: rise_logo_quarto.png
    mainfont: 'Lato'
    monofont: 'Roboto Mono'
    code-overflow: wrap
    code-tools: true
    code-fold: true
    number-sections: true
    fig-dpi: 300
    layout-align: left
  pdf:
    papersize: a4
    documentclass: article #article, report or book
    classoption: [twocolumn, portrait]
  revealjs:
    theme: default
    logo: rise_logo_quarto.png
    chalkboard: false
    self-contained: true
    footer: 'Created by magnus.p.johansson@ri.se'
    mainfont: 'Lato'
    slide-level: 4
    scrollable: true
    smaller: false
  docx:
    toc: true
    number-sections: true
    title: "Rasch in Quarto"
    subtitle: "Template file"
    author: "Magnus Johansson"
#    reference-doc: RISEmallMJv6.dotx
execute:
  echo: false
  warning: false
  message: false
  cache: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: inline
---

```{r}
#| label: setup
#| code-fold: false
#| include: false

# two packages below require that you use devtools to install them manually:
# first install devtools by
# install.packages('devtools')

library(ggrepel)
library(car)
library(grateful) # devtools::install_github("Pakillo/grateful")
library(kableExtra)
library(readxl)
library(tidyverse)
library(eRm)
library(mokken)
library(mirt)
library(psych)
library(ggplot2)
library(psychotree)
library(matrixStats)
library(reshape)
library(knitr)
library(cowplot)
library(formattable) 
library(RISEkbmRasch) # devtools::install_github("pgmj/RISEkbmRasch")
library(HH)

### some commands exist in multiple packages, here we define preferred ones that are frequently used
select <- dplyr::select
count <- dplyr::count
recode <- car::recode
rename <- dplyr::rename

### set up color palette based on RISE guidelines
RISEprimGreen <- "#009ca6"
RISEprimRed <- "#e83c63"
RISEprimYellow <- "#ffe500"
RISEprimGreenMid <- "#8dc8c7"
RISEprimRedMid <- "#f5a9ab"
RISEprimYellowMid <- "#ffee8d"
RISEprimGreenLight <- "#ebf5f0"
RISEprimRedLight <- "#fde8df"
RISEprimYellowLight <- "#fff7dd"
RISEcompPurple <- "#482d55"
RISEcompGreenDark <- "#0e4e65"
RISEgrey1 <- "#f0f0f0"
RISEgrey2 <- "#c8c8c8"
RISEgrey3 <- "#828282"
RISEgrey4 <- "#555555"

# set some colors used later
cutoff_line <- RISEprimRed
dot_color <- "black"
backg_color <- RISEprimGreenLight

# set fontsize for all tables
r.fontsize <- 15

### first we pre-set our chosen cut-off values for some commonly used indices:
msq_min <- 0.7
msq_max <- 1.3
zstd_min <- -2
zstd_max <- 2
loc_dep <- 0.2 # above average residual correlation
dif_dif <- 0.5 # logits difference between groups in average item location (DIF)

### zstd is inflated with large samples (N > 500). Reduce sample size to jz and 
### run analysis yz random samples to get average ZSTD
jz = 300 # number to include in dataset
yz = 10 # number of random samples

#---- Load data----
df<-as.data.frame(read_excel("GNI23data v1.1.xls"))

### if you need to download the datafile:
# library(readODS)
# url <- "https://figshare.com/ndownloader/files/27953790"
# destfile <- "gnidata.ods"
# download.file(url, destfile, mode="wb")
# df <- read_ods(file = "gnidata.ods")

# Load item information
itemlabels<-read_excel("WAAQitemlabels.xls")
#responseOptions<-read_excel("/Users/magnuspjo/Library/CloudStorage/OneDrive-SharedLibraries-RISE/KP Centrum för Kategoriskt Baserade Mätningar - Documents/Kunskapsutveckling/RaschR/data/PSS10itemlabels.xls", sheet = 2)
#introText<-read_excel("/Users/magnuspjo/Library/CloudStorage/OneDrive-SharedLibraries-RISE/KP Centrum för Kategoriskt Baserade Mätningar - Documents/Kunskapsutveckling/RaschR/data/PSS10itemlabels.xls", sheet = 3)

# Make a backup of the dataframe
df.all<-df

#---- Create a vector with variables names----
scale.items <- itemlabels$itemnr

# If desired, you can instead create the scale.items object from the dataframe variable names:
scale.items <- df %>%
  select(starts_with("WAAQ")) %>% # this selects all columns/variables that start with WAAQ
  names()

#---- recode responses to WAAQ items to 0-6 (instead of 1-7)----
for (i in scale.items) {
  df[[i]]<-recode(df[[i]],"1=0;2=1;3=2;4=3;5=4;6=5;7=6",as.factor=FALSE)
}


#---- Create a figure showing % of missing data for each item, based on the complete dataset----
missingdata <- df %>% 
  select(starts_with("WAAQ")) %>% # here you select which items are to be shown in the figure
  t() %>% 
  as.data.frame() %>% 
  mutate(Missing = rowSums(is.na(.))) %>% 
  select(Missing) %>% 
  arrange(desc(Missing)) %>% 
  rownames_to_column(var = "Item") %>% 
  mutate(Percentage = Missing/nrow(df)*100) %>% 
  mutate(Item = factor(Item, levels = rev(Item))) %>%
  ggplot(aes(x = Item, y = Percentage)) +
  geom_col() +
  coord_flip() +
  ggtitle("Missing data per item") +
  xlab("Items") +
  ylab("Percentage of responses missing")

# You can display the figure created above anywhere in the Quarto document by inserting the code `r missingdata` in text areas (with the `` symbols intact), or by just writing "missingdata"on a single line in a code chunk (without the "" symbols).

#---- Filtering participants based on missing data----

# If you want to include participants with missing data, input the minimum number of items responses that a participant should have to be included in the analysis:
min.responses <- 3

# Select the variables we will work with, and filter out respondents with a lot of missing data
df.omit.na <- df %>% 
  select(starts_with("WAAQ"),Sex,Age) %>%  # variables that start with "WAAQ"
  filter(length(scale.items)-rowSums(is.na(.[scale.items])) >= min.responses) # include only respondents with data for at least 3 items

#---- OR just filter out all respondents with any missing data----
df.omit.na <- df %>% 
  select(starts_with("WAAQ"),Sex,Age) %>% 
  na.omit()

#---- Create DIF variables----
  
# DIF variables into vectors
dif.gender <- df.omit.na$Sex
dif.age <- df.omit.na$Age
# then remove them from dataframe, since we need a dataframe with only item data for the Rasch analyses
df.omit.na$Sex <- NULL
df.omit.na$Age <- NULL

##### If desired, set item names to brief and neutral "qX" (and avoid issues with the character "_" in pdf output)
##### The important thing is that variable names in the data match the variablenames in the itemlabels dataframe
#names(df.omit.na) <- paste0("q", c(1:ncol(df.omit.na)))

```

## All items in the analysis
```{r}
RIlistitems(df.omit.na)
```

## Demographics
```{r}
#| layout-ncol: 2
RIdemographics(dif.gender,"Kön")
# if you want to recode a numerical variable to text for the table (or vice versa), you can use this line
#dplyr::recode(dif.gender, '1' = "Male", '2' = "Female") %>% RIdemographics("Gender")
RIdemographics(dif.age,"Ålder")
```

### Descriptives of raw data

Responses to all items are summarized below.

```{r}
#| tbl-cap: "Total number of responses for all items"
RIallresp(df.omit.na)
```

### Descriptives - item level

::: panel-tabset
#### Tile plot
```{r}
RItileplot(df.omit.na)
```
#### Stacked bars
```{r}
RIbarstack(df.omit.na)
```
#### Barplots
```{r}
#| layout-ncol: 2
RIbarplot(df.omit.na)
```
:::

### Floor/ceiling effects - raw data {.smaller}
```{r}
RIrawdist(df.omit.na)
```

## Analysis of response categories {.smaller}

```{r}
#| include: false
mirt.rasch <- mirt(df.omit.na, model=1, itemtype='Rasch') # unidimensional Rasch model
```
```{r}
plot(mirt.rasch, type="trace")
```

## Rasch analysis 1 {.smaller}

The eRm package, which uses Conditional Maximum Likelihood (CML)
estimation, will be used primarily. For this analysis, the Partial
Credit Model will be used.

**The section below is useful to copy and re-run after making changes to the data, such as removing ill-fitting items or recoding disordered response categories.**

::: panel-tabset
### Item fit {.smaller}
```{r}
RIitemfitPCM(df.omit.na)
```
### PCA
```{r}
#| tbl-cap: "PCA of Rasch model residuals"
RIpcmPCA(df.omit.na)
```
### Residual correlations {.smaller}
```{r}
RIresidcorr(df.omit.na, cutoff = 0.2)
```
## 1st contrast loadings
```{r}
RIloadLoc(df.omit.na)
```
### Targeting
```{r}
RItargeting(df.omit.na)
```
### Targeting sorterad
```{r}
df.erm<-PCM(df.omit.na)
plotPImap(df.erm, sorted = T)
```
### Items
```{r}
RIlistitems(df.omit.na)
```
:::

#### Item information {.smaller}
"Item information" shows individual curves indicating the amount of
information contributed by each item.
```{r}
df.erm <- PCM(df.omit.na)
plotINFO(df.erm, type = "item", legpos = "topleft") # get item information curves
```

### Person location and infit ZSTD {.smaller}
```{r}
RIpfit(df.omit.na)
```

## Item parameters
```{r}
RIitemparams(df.omit.na)
```

## DIF-analysis

::: panel-tabset
### Tabell
```{r}
RIdifTable(df.omit.na, dif.gender)
```
### Figur
```{r}
RIdifFigure(df.omit.na, dif.gender)
```
### Items
```{r}
RIlistitems(df.omit.na)
```
:::

## Reliability
```{r}
RItif(df.omit.na)
```

## Item location & infit ZSTD
```{r}
RIinfitLoc(df.omit.na)
```

## Item location & outfit ZSTD
```{r}
RIoutfitLoc(df.omit.na)
```

## Software used
```{r}
pkgs <- cite_packages(cite.tidyverse = TRUE, 
                      output = "table",
                      bib.file = "grateful-refs.bib",
                      include.RStudio = TRUE)
formattable(pkgs, 
            table.attr = 'class=\"table table-striped\" style="font-size: 15px; font-family: Lato; width: 80%"')

```

